Fourier-domain modulations and delays of gravitational-wave signals

### 2017-06-07 docs/paper_method.tex

A useful representation of the gravitational waveform is given by its decomposition in spin-weighted spherical. The waveform emitted in the direction $(\Theta, \Phi)$, with its two polarizations $h_{+},h_{\times}$, can be decomposed as a superposition of modes as~\cite{Thorne80}
\be\label{eq:defmodes}
	h_{+} - i h_{\times} = \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} {}_{-2}Y_{\ell m}(\Theta,\Phi) h_{\ell m} \,,
\ee
where the ${}_{-2}Y_{\ell m}(\Theta,\Phi)$ are spin-weighted spherical harmonics. For a non-precessing system, with a fixed orbital plane, the individual modes have the additional symmetry property $h_{\ell -m} = (-1)^{\ell} h_{\ell m}^{*}$. Since we will work in the Fourier domain, it will be useful to write the contributions of the individual modes to the Fourier transforms of the polarizations as
\begin{subequations}\label{eq:hpcfrommodes}
\begin{align}
	\tilde{h}_{+}(f) &= \frac{1}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left( {}_{-2}Y_{\ell m} \widetilde{h_{\ell m}}(f) + {}_{-2}Y_{\ell m}^{*} \widetilde{h_{\ell m}}(-f)^{*} \right) \,, \\
	\tilde{h}_{\times}(f) &= \frac{i}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left( {}_{-2}Y_{\ell m} \widetilde{h_{\ell m}}(f) - {}_{-2}Y_{\ell m}^{*} \widetilde{h_{\ell m}}(-f)^{*} \right) \,,
\end{align}
\end{subequations}
where we used $\widetilde{h_{\ell m}^{*}}(f) = \widetilde{h_{\ell m}}(-f)^{*}$. An additional approximation often used in waveform models consists in considering that the Fourier transforms $\tilde{h}_{\ell m}$ have support only on one side of the spectrum, either for positive of for negative frequencies depending on the sign of $m$. This holds in particular within the stationary phase approximation, as explained below. With our sign convention~\eqref{eq:defFT}, this approximation reads
\begin{align}\label{eq:approxnegativef}
	\tilde{h}_{\ell m} (f) &\simeq 0 \text{ for } f<0, \; m>0 \nn\,,\\
	\tilde{h}_{\ell m} (f) &\simeq 0 \text{ for } f>0, \; m<0 \,.
\end{align}

@@

In this Section, we focus on the second-derivative corrective term in~\eqref{}. We will restrict to the case of a pure modulation, $d=0$ and $G(f,t) = F(t)$ [go back to more general $G(f,t)$ ?]. The authors of~\cite{KCY14} proposed a resummation scheme. Indeed, the result~\eqref{} looks like a symmetrized Taylor expansion, except for the factors $i^{p}$ and $1/p!$ instead of $1/(2p)!$. After having truncated the sum at some finite $N$, one can write (following~\cite{KCY14})
\begin{align}\label{eq:resumquadphase}
	\calT(f) &\simeq \sum\limits_{p = 0}^{N} \frac{(-i\epsilon\Tf^{2})^{p}}{2^{p}p!} \partial_{t}^{2p}F(\tf) \nn\\
	&\simeq \sum\limits_{p= 0}^{N} \sum\limits_{k=0}^{N} a_{N,k}\frac{(k\Tf)^{2p}}{(2p)!}  \partial_{t}^{2p}F(\tf) \nn\\
	&\simeq \frac{1}{2}\sum\limits_{k=0}^{N} a_{N,k} \left( F(\tf + k\Tf) + F(\tf-k\Tf) \right)\,,
\end{align}
where one applies a truncation approximation at the first and last stage. The second equality holds provided that the complex coefficients $a_{N,k}$ are a solution of the $N+1$-dimensional linear system
\be\label{eq:stencilsystem}
	(-i\epsilon)^{p} (2p-1)!! = \sum\limits_{k=0}^{N} a_{N,k} k^{2p} \quad \text{for } p=0,\dots,N \,.
\ee
The two solutions for the stencil coefficients in the two cases $\epsilon = \pm 1$ are simply related by a complex conjugation.

>>>>

A useful representation of the gravitational waveform is given by its decomposition in spin-weighted spherical harmonics. The waveform emitted in the direction $(\Theta, \Phi)$, with its two polarizations $h_{+},h_{\times}$, can be decomposed as a superposition of modes as~\cite{Thorne80}
\be\label{eq:defmodes}
	h_{+} - i h_{\times} = \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} {}_{-2}Y_{\ell m}(\Theta,\Phi) h_{\ell m} \,,
\ee
where the ${}_{-2}Y_{\ell m}(\Theta,\Phi)$ are spin-weighted spherical harmonics. For a non-precessing system, with a fixed orbital plane, the individual modes have the additional symmetry property $h_{\ell -m} = (-1)^{\ell} h_{\ell m}^{*}$. Since we will work in the Fourier domain, it will be useful to write the contributions of the individual modes to the Fourier transforms of the polarizations as
\begin{subequations}\label{eq:hpcfrommodes}
\begin{align}
	\tilde{h}_{+}(f) &= \frac{1}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left( {}_{-2}Y_{\ell m} \widetilde{h_{\ell m}}(f) + {}_{-2}Y_{\ell m}^{*} \widetilde{h_{\ell m}}(-f)^{*} \right) \,, \\
	\tilde{h}_{\times}(f) &= \frac{i}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left( {}_{-2}Y_{\ell m} \widetilde{h_{\ell m}}(f) - {}_{-2}Y_{\ell m}^{*} \widetilde{h_{\ell m}}(-f)^{*} \right) \,,
\end{align}
\end{subequations}
where we used $\widetilde{h_{\ell m}^{*}}(f) = \widetilde{h_{\ell m}}(-f)^{*}$. An additional approximation often used in waveform models consists in considering that the Fourier transforms $\tilde{h}_{\ell m}$ have support only on one side of the spectrum, either for positive of for negative frequencies depending on the sign of $m$. This holds in particular within the stationary phase approximation, as explained below. With our sign convention~\eqref{eq:defFT}, this approximation reads
\begin{align}\label{eq:approxnegativef}
	\tilde{h}_{\ell m} (f) &\simeq 0 \text{ for } f<0, \; m>0 \nn\,,\\
	\tilde{h}_{\ell m} (f) &\simeq 0 \text{ for } f>0, \; m<0 \,.
\end{align}

@@

In this Section, we focus on the second-derivative corrective term in~\eqref{}. We will restrict to the case of a pure modulation, $d=0$ and $G(f,t) = F(t)$ [go back to more general $G(f,t)$ ?]. The authors of~\cite{KCY14} proposed a resummation scheme. Indeed, the result~\eqref{} looks like a symmetrized Taylor expansion, except for the factors $i^{p}$ and $1/p!$ instead of $1/(2p)!$. After having truncated the sum at some finite $N$, one can write (following~\cite{KCY14})
\begin{align}\label{eq:resumquadphase}
	\calT(f) &\simeq \sum\limits_{p = 0}^{N} \frac{(-i\epsilon\Tf^{2})^{p}}{2^{p}p!} \partial_{t}^{2p}F(\tf) \nn\\
	&\simeq \sum\limits_{p= 0}^{N} \sum\limits_{k=0}^{N} a_{N,k}\frac{(k\Tf)^{2p}}{(2p)!}  \partial_{t}^{2p}F(\tf) \nn\\
	&\simeq \frac{1}{2}\sum\limits_{k=0}^{N} a_{N,k} \left( F(\tf + k\Tf) + F(\tf-k\Tf) \right)\,,
\end{align}
where one applies a truncation approximation at the first and last stage. The second equality holds provided that the complex coefficients $a_{N,k}$ are a solution of the $N+1$-dimensional linear system~\cite{KCY14}
\be\label{eq:stencilsystem}
	(-i\epsilon)^{p} (2p-1)!! = \sum\limits_{k=0}^{N} a_{N,k} k^{2p} \quad \text{for } p=0,\dots,N \,.
\ee
The two solutions for the stencil coefficients in the two cases $\epsilon = \pm 1$ are simply related by a complex conjugation.
